---
title: "Word Prediction"
author: "Pep Carrera"
date: "November 18, 2014"
output: html_document
runtime: shiny
---

This application takes a sentance you type in and standardizes the last three words (setting them to lower case, removing special characters, etc).  It uses the last three words of your sentance to give you a set of possible next three words, thereby prediciting your next word.  If you only type in one word, it uses just one, followed by two, and up to three. 

This is accomplished using a corpus of text from twitter, blogs, and news articles.

```{r, cache=TRUE, echo=FALSE}
library(stringr)

standardizeWord <- function(word) {
        word <- str_replace_all(word, "=|!|,|/.|-|_|#|?|!|'|:|*", "")
        word <- tolower(word)
        return(word)
}

predictWord <- function(word1, word2="", word3="") {
        if (word3 == "" & word2 == "") {
                word1 <- standardizeWord(word1)
                predDf2 <- twoFreq[with(twoFreq, w1 == word1),]
                predWord1 <- toString(unlist(predDf2[1,2]))
                predWord2 <- toString(unlist(predDf2[2,2]))
                predWord3 <- toString(unlist(predDf2[3,2]))
        } else if (word3 == "") {
                word1 <- standardizeWord(word1)
                word2 <- standardizeWord(word2)
                predDf2 <- twoFreq[with(twoFreq, w1 == word2),]
                predDf3 <- threeFreq[with(threeFreq, w1 == word1 & w2 == word2),]
                predWord1 <- toString(unlist(predDf3[1,3]))
                predWord2 <- toString(unlist(predDf3[2,3]))
                predWord3 <- toString(unlist(predDf3[3,3]))
                if (predWord1 == "NA") {
                        predWord1 <- toString(unlist(predDf2[1,2]))
                        predWord2 <- toString(unlist(predDf2[2,2]))
                        predWord3 <- toString(unlist(predDf2[3,2]))
                } else if (predWord2 == "NA" & predWord3 == "NA") {
                        predWord2 <- toString(unlist(predDf2[2,2]))
                        predWord3 <- toString(unlist(predDf2[3,2]))  
                } else if (predWord3 == "NA") {
                        predWord3 <- toString(unlist(predDf2[3,2]))
                }
        } else {
                word1 <- standardizeWord(word1)
                word2 <- standardizeWord(word2)
                word3 <- standardizeWord(word3)
                predDf2 <- twoFreq[with(twoFreq, w1 == word3),]
                predDf3 <- threeFreq[with(threeFreq, w1 == word2 & w2 == word3),]
                predDf4 <- fourFreq[with(fourFreq, w1 == word1 & w2 == word2 & w3 ==word3),]
                predWord1 <- toString(unlist(predDf4[1,4]))
                predWord2 <- toString(unlist(predDf4[2,4]))
                predWord3 <- toString(unlist(predDf4[3,4]))
                if (predWord1 == "NA") {
                        predWord1 <- toString(unlist(predDf3[1,3]))
                        predWord2 <- toString(unlist(predDf3[2,3]))
                        predWord3 <- toString(unlist(predDf3[3,3]))
                } else if (predWord2 == "NA" & predWord3 == "NA") {
                        predWord2 <- toString(unlist(predDf3[2,3]))
                        predWord3 <- toString(unlist(predDf3[3,3]))  
                } else if (predWord3 == "NA") {
                        predWord3 <- toString(unlist(predDf3[3,3]))
                }
                
                if (predWord1 == "NA") {
                        predWord1 <- toString(unlist(predDf2[1,2]))
                        predWord2 <- toString(unlist(predDf2[2,2]))
                        predWord3 <- toString(unlist(predDf2[3,2]))
                } else if (predWord2 == "NA" & predWord3 == "NA") {
                        predWord2 <- toString(unlist(predDf2[2,2]))
                        predWord3 <- toString(unlist(predDf2[3,2]))  
                } else if (predWord3 == "NA") {
                        predWord3 <- toString(unlist(predDf2[3,2]))
                }
        }
        
        checkWord(predWord1, predWord2, predWord3)
}

checkWord <- function(predWord1, predWord2, predWord3) {
        if (predWord1 == "NA") predWord1 <- noWord[1]
        if (predWord2 == "NA") predWord2 <- noWord[2]
        if (predWord3 == "NA") predWord3 <- noWord[3]
        predWords <- c(predWord1, predWord2, predWord3)
        return(predWords)
}

inputConversion <- function(sent) {
lastThree <- standardizeWord(tail(strsplit(sent,split=" ")[[1]],3))
}

twoFreq <- readRDS("data/twoFreq.rds")
threeFreq <- readRDS("data/threeFreq.rds")
fourFreq <- readRDS("data/fourFreq.rds")
noWord <- readRDS("data/noWord.rds")
```

```{r, echo=FALSE}
textInput('sent', 'Type a sentance:')

renderText({
lastThree <- inputConversion(input$sent)
paste("First Word Used to Predict is:", lastThree[1])
})

renderText({
lastThree <- inputConversion(input$sent)
paste("Second Word Used to Predict is:", lastThree[2])
})

renderText({
lastThree <- inputConversion(input$sent)
paste("Third Word Used to Predict is:", lastThree[3])
})

```


Your predicted words are below, if it shows NULL, type something!
```{r, echo=FALSE}

renderPrint({
lastThree <- character()
#Call input conversion so we can get the lats three words (and standadize them)
lastThree <- inputConversion(input$sent)

# call the prediction function with the appropraite arguments
if (length(lastThree) == 0) {
} else if (length(lastThree) == 1) {
  predictWord(word1=lastThree[1])
} else if(length(lastThree) == 2) {
  predictWord(word1=lastThree[1],word2=lastThree[2])
} else {
  predictWord(word1=lastThree[1],word2=lastThree[2], word3=lastThree[3])
}

})
```



